Dancing on Water: The Choreography of Sulfur Dioxide Adsorption to Aqueous Surfaces

### 2011-10-25 resubmission/ReviewResponses.tex

The authors greatly appreciate the detail of the reviews provided, and have made significant changes to the manuscript to incorporate the suggestions, and to clarify the various points raised. Below are responses to each of the issues raised by the reviewers.

\subsection*{Review 2}

\begin{enumerate}

\item The steered MD technique used in our study is one established previously in the publications referenced in the introduction of the manuscript (26-31). Specifically, the theoretical foundation of the technique was established by Isralewitz (Current opinion in structural biology [0959-440X] Isralewitz yr:2001 vol:11 iss:2 pg:224-230). The technique applies a constant force or velocity to any combination of atoms or atomic subgroups in the system in order to affect a motion. In our SMD simulations, the amount of force applied is very small compared to the intermolecular forces of the aqueous surface, but the applied force is constant, and slowly moves the transiting \suldiox~into the aqueous surface from the gas. We believe the method to be sound, and have experimented with several of the parameters to find a combination that does not artificially bias the geometries and orientations of the atoms involved.

\item In this work we attempted to two distinct studies that involve bulk-solvated \suldiox~as a starting configuration. The first study looks at a single \suldiox~already adsorbed to a water surface (the neat-water system). The aqueous system was then equilibrated with \suldiox~already bound and solvated. Evolution of the system through an extended equilibration resulted in a very surface-active \suldiox, which is the expected result both from our own surface-specific experiments, and from the results of Baer et al, that simulate this system identically (save for the choice of water model used). Thus the initial artificially solvated \suldiox~evolved to a natural location near the aqueous surface following equilibration.

The second study (i.e. the "saturated" system) used the solvated \suldiox~as a starting configuration as well, but with two distinctions from the "neat-water" system. First there were several more \suldiox~molecules in the bulk instead of a single molecule. Secondly, \suldiox~was introduced in the gas-phase as well, and subsequently adsorbed to the water surface during the equilibration phase. Although the starting conditions of bulk solvation in both the neat-water and saturated systems were not necessarily representative of the \suldiox~behavior, the extended equilibration phase of our simulations evolved the \suldiox~to a location and system geometry that is consistent with our experimental results (surface active), and with the results of simulations carried out previously by Baer et al (residual sub-surface \suldiox). Thus, our "surface equilibrated method" finds justification in previous published results.

Regarding the steered MD (SMD), the systems were established and equilibrated in a manner identical to the surface equilibrated method. Although the starting point of the equilibration may not have been representative of an equilibrium system of solvated \suldiox, the configurations at the start of our data collection were what we expected, and what previous studies dictated would occur in terms of \suldiox~location and behavior. The subsequent SMD then introduced \suldiox~from the gas phase, which does not introduce into the system artificial behavior, as \suldiox~gas is known to adsorb from the vapor (as pointed out in the suggested literature of Donaldson et al, below, and from our own experimental studies).

@@

We believe that our introduction to the manuscript specifically addresses what is and is not known, and how the present work adds to our knowledge of sulfur dioxide behavior in aqueous/vapor interfacial environments. We emphasize here again that our study was not set to be a review of the field of gaseous uptake, but a narrowly defined study of sulfur dioxide orientational behavior on an aqueous surface, and a complement to our experimental studies on the same system. Thus we did not include a discussion relating to other gaseous molecules or related systems. Those discussions have already been published in several of our previous experimental studies that we reference in the manuscript.

\item The suggestion to use bi-variate distributions is appreciated and well made. We have significantly reanalyzed our data sets and presented the relevant depth-profiles as bi-variate plots P($\theta$, $\phi$) showing the specific orientations of water and sulfur dioxide at various depths (slices) throughout the interfacial region. Significant changes have been made to the text to accommodate the new plots.

@@

\item The "Computational Approach" section does not contain any text regarding results or analysis. The section is reserved to define the computational approach used, and to define the terminology used regarding depth and orientational analysis.

\item The term "absorption" is used only once in the computational methods section. Regarding the abstract, we have submitted a corrected version and greatly appreciate mention of this error.

\item To address the used of the POL3 model with the \suldiox~model developed by Baer et al, we have submitted a supplemental analysis to be included with the manuscript. We have calculated the radial distribution functions of the \wat-\suldiox~system to compare the resultant structure and geometries of the interactions between the two types of molecules. Furthermore, we have calculated the \suldiox-\wat~interaction as a function of the S-O separation. The results of both of these analyses are directly comparable to the results of Baer et al. Our calculations show a nearly identical result to Baer's in both the interaction energy and liquid structure using the POL3 water model. Thus we feel this justifies the use of the more computationally efficient POL3 model.

\item The Henry's law constant for \suldiox~in water has been added to the text.

\item The pressure in the gas region was set initially to 1 atm. The system was then equilibrated until all the gaseous \suldiox~adsorbed to the water surface, at which point another 1 atm of \suldiox~was introduced. This procedure was repeated until the gas phase \suldiox~no longer adsorbed to the surface. This required a total of 50 \suldiox~molecules introduced in the gas phase. Thus a constant pressure of 1 atm was used throughout equilibration and data collection simulations. We do not feel that altering the text is a necessary step regarding this point as the bulk of \suldiox~molecules adsorb to the surface and do not remain in gas to contribute to the overall gas pressure. A sentence regarding this was added to the methods section to clarify the procedure.

\item We have addressed the technical details of the simulation conditions in the text of the manuscript.

\item Two methods sections, one each for the surface equilibrated MD and the steered MD, were specifically included to resolve any doubt regarding the simulation conditions of the two different techniques used. The text may appear to be redundant, but we are addressing two separate simulation techniques.

@@

\item The terminology (e.g. "surface equilibrated method") has been chosen as the most appropriate descriptor for the system to clearly distinguish the equilibrated system from the steered MD system. System descriptors have been used consistently throughout the manuscript, and we feel that they are appropriate to the current work.

\item 
\begin{enumerate}
\item All simulations were performed using the NSCM flag in Amber, and the surface location drift still occurs. This is not a drift of the entire system center of mass, but a slight swelling or movement of the slab over time. The problem is not remedied by the suggested solution.

### 2011-10-25 Conclusions.tex
Final resubmission to the journal

Our simulations show that gaseous \suldiox~quickly adsorbs to the water surface and continues to bind until a complete surface coverage is reached. Surface waters reorient in the presence of adsorbed \suldiox. The waters at the interface of a neat-water surface tend to lay flatter to the surface than when a saturating layer of \suldiox~is present. The waters interacting with the layer of adsorbed \suldiox~orient more perpendicularly to the interface, and further expose their ``free-OH'' uncoupled bonds for interactions with \suldiox, and hydrate complex formation. Furthermore, we have found that surface waters underneath a blanket layer of adsorbed \suldiox~will penetrate further into the gas phase, allowing for greater mobility of waters away from the aqueous bulk in the presence of \suldiox.

Through these simulations we also characterize the orientational behavior of \suldiox~during and after adsorption. Our equilibrium neat-water simulations show that a single \suldiox~molecule, representing a low concentration, has a high surface affinity. At a high \suldiox~concentration in the saturated systems, \suldiox~molecules are also surface active, and are found further out of the water phase than at the lower concentration. These \suldiox~molecules form a bound layer that crowds the surface and interacts with the surface waters. The orientation of \suldiox~on the water surface was found to be similar for both low and high concentrations. Those \suldiox~molecules at or below the surface water location strongly orient with the sulfur atom pointed in towards the water bulk, and the oxygen atoms out towards the gas phase. The \suldiox~molecules slightly above the water surface lose this net orientation within 10 \angs. Those  molecules further from the water are more isotropically oriented. Figure \ref{fig:so2-surface-cartoon} depicts what the neat-water and saturated surface molecules look like for both \suldiox~and \wat~orientations and locations based on the calculations.

@@

Steered molecular dynamics simulations were used to model the behavior of an adsorbing \suldiox~as it moves from the gas phase above the water down through the surface and into the bulk. The \suldiox~reorients as it makes its first contact with the water interface. Within 5 \angs~of the surface the \suldiox~is mostly oriented with its sulfur towards the water phase. The results for the transit through the interface show that in both systems of low and high \suldiox~concentration an adsorbing \suldiox~near the interface has very similar orientation to those molecules already bound to the water surface. The \suldiox~pulled further into the water bulk retains its orientation until it is past the interfacial region and then isotropically orients with the bulk water.

>>>>

Our simulations show that gaseous \suldiox~quickly adsorbs to the water surface and continues to bind until a complete surface coverage is reached. Surface waters reorient in the presence of adsorbed \suldiox. The waters at and just below the interface of a neat-water surface tend to lay flatter to the surface than when a saturating layer of \suldiox~is present. The waters above the surface location or interacting with the layer of adsorbed \suldiox~orient more perpendicularly to the interface, and further expose their ``free-OH'' uncoupled bonds for interactions with \suldiox, and hydrate complex formation. Furthermore, we have found that surface waters underneath a blanket layer of adsorbed \suldiox~will penetrate further into the gas phase, allowing for greater mobility of waters away from the aqueous bulk in the presence of \suldiox.

Through these simulations we also characterize the orientational behavior of \suldiox~during and after adsorption. Our equilibrium neat-water simulations show that a single \suldiox~molecule, representing a low concentration, has a high surface affinity. At a high \suldiox~concentration in the saturated systems, \suldiox~molecules are also surface active, and are found further out of the water phase than at the lower concentration. These \suldiox~molecules form a bound layer that crowds the surface and interacts with the surface waters. The orientation of \suldiox~on the water surface was found to be similar for both low and high concentrations. Those \suldiox~molecules at or below the surface water location strongly orient with the sulfur atom pointed in towards the water bulk, and the oxygen atoms out towards the gas phase. The \suldiox~molecules slightly above the water surface lose this net orientation within 6-8 \angs. Those  molecules further from the water are more isotropically oriented. Figure \ref{fig:so2-surface-cartoon} depicts what the neat-water and saturated surface molecules look like for both \suldiox~and \wat~orientations and locations based on the calculations.

@@

Steered molecular dynamics simulations were used to model the behavior of an adsorbing \suldiox~as it moves from the gas phase above the water down through the surface and into the bulk. The \suldiox~reorients as it makes its first contact with the water interface. Within 4 \angs~of the surface the \suldiox~is mostly oriented with its sulfur towards the water phase. The results for the transit through the interface show that in both systems of low and high \suldiox~concentration an adsorbing \suldiox~near the interface has very similar orientation to those molecules already bound to the water surface. The \suldiox~pulled further into the water bulk retains its orientation until it is past the interfacial region and then isotropically orients with the bulk water.

### 2011-10-25 EquilibratedOrientation.tex
Final resubmission to the journal

Geometric analyses were performed to characterize the net molecular orientation of \wat~and \suldiox~molecules at different depths from the water surface location. At each distance from the surface location, an orientation profile was created for both the \wat~and \suldiox~molecules. The bi-variate orientation distributions for the angles $\theta$ and $\phi$ at various depths were combined to form the intensity plots that show how the molecular orientation distributions change with distance to the surface location. These plots (Figures \ref{fig:water-orientation}, \ref{fig:so2-transit-angles}, and \ref{fig:so2-surface-angles}) allow for a visual interpretation of how the net orientations are affected when moving from the gas phase through the interfacial region and to the surface location, and then further into the aqueous interfacial region and bulk. Both the neat-water, with only a single \suldiox~introduced, and the high-concentration saturated system were analyzed. In the case of the neat-water system, the introduction of a single \suldiox~does not greatly affect molecular orientation of water molecules in the interfacial region. These results of the water orientation are very similar to a neat-water system without any adsorbed solutes (not shown).

@@

Although the plots show overall similarities for both the neat-water and saturated systems, the presence of a layer of adsorbed \suldiox~molecules alters the orientation of those waters furthest into the gas phase. For the saturated solution, the resulting orientation of waters above 0\angs, shown in the bottom plot of Figure \ref{fig:water-orientation}, is nearly isotropic in $\phi$, and with $\theta leq$ 90\textdegree. This results from waters with bisectors pointing further into the adsorbed \suldiox~gas layer, and both hydrogens pointing outward from the aqueous bulk. The effect is more pronounced as the waters move further from the water surface, and above 4\angs~the $\theta$ distribution is mostly centered around $\theta)=$ 0\textdegree (see Figure \ref{fig:water-angles}).

@@

The distribution of $\cos(\phi)$ is more sharply defined (i.e. less isotropic) for the neat-water system than for the saturated one. Waters on the neat surface lie flatter, whereas the presence of the \suldiox~allows a greater range of ``twist'' for those waters in the plane of the interface. The $\phi$ distributions quickly become isotropic above the surface location as the bisectors orient more perpendicularly, and below the surface as the bulk water loses any orientational preference.

@@

Orientation distributions of the adsorbed \suldiox~molecules were created during the equilibrium simulations for both the neat-water and saturated systems. Figure \ref{fig:so2-orientation} shows the 2D distributions of $\theta$ and $\phi$ (arranged similarly to the water orientation distributions plots in Figure \ref{fig:water-orientation}). The \suldiox~orientation data set for the neat-water system is much smaller as only a single \suldiox~molecule was simulated in the bulk. The resulting distribution plots are thus representative of the single surface active \suldiox~molecule. The neat-water \suldiox~molecule remains within a narrower region of the interface than the saturated system \suldiox, but effective comparisons can still be drawn.

In the interfacial region the angular distribution of the single \suldiox~(in the neat-water system) is concentrated primarily in $\theta <$ 90 \textdegree. The peak of the distribution occurs at $\theta)=$ 0\textdegree. This indicates that the \suldiox~bisector points out of the water surface, with the sulfur atom pointing towards the aqueous bulk, and the two oxygens pointing into the gas phase. This same distribution occurs in the saturated system for depths below the surface location, $< 0$\angs. Beyond 4\angs~above the surface, both distributions become mostly isotropic, either because the \suldiox~does not venture into the gas in the neat-water system, or because of the nature of the adsorbed \suldiox~layer in the saturated system. Promixity to the water surface highly orients the \suldiox~bisector. % with the sulfur atom pointing in towards the water bulk. %Moving further away from the water surface, and interacting less with \wat~molecules allows for greater orientational freedom as exhibited in the isotropy of the distributions above 4\angs. %Because both the neat-water and saturated systems show rather similar distributions, it is possible that the concentration of \suldiox~does not strongly affect the orientation of the \suldiox, unlike the orientations of waters.

The distributions of $\phi$ are isotropic in both systems at all depths. Because the \suldiox~bisector near the surface is oriented perpendicularly to the interface, the isotropy in $\phi$ is expected. Further from the water surface where the bisector orientation becomes isotropic, the $\cos(\phi)$ distribution remains isotropic. For the \suldiox~orientation, the $\phi$ angle does not provide further information regarding the surface behavior or orientational preference.

@@

\begin{figure}[h!]
	\begin{center}
		\includegraphics[scale=1.0]{images/so2-angles/theta-phi.png}
		\caption{Molecular orientation distributions for \suldiox~molecules adsorbed to the water slab surface. The angular distributions are shown for the neat-water (top) and saturated (bottom) systems, and are arranged similarly to the water orientation plots of Figure \ref{fig:water-orientation}. For both systems the distributions show \suldiox~bound to the water surface with the sulfur primarily pointing towards the water slab, and the oxygens pointing to the gas phase. In this configuration the $\phi$ distribution is isotropic because of the water slab's in-plane symmetry. Note that the range of surface depths in the low concentration system is half that of the saturated system. This is due to the narrower position range of the single \suldiox~molecule.}
		\label{fig:so2-orientation}
	\end{center}
\end{figure}

>>>>

Geometric analyses were performed to characterize the net molecular orientation of \wat~and \suldiox~molecules at different depths from the water surface location. At each distance from the surface location, an orientation profile was created for both the \wat~and \suldiox~molecules. The bi-variate orientation distributions for the angles $\theta$ and $\phi$ at various depths were combined to form the intensity plots that show how the molecular orientation distributions change with distance to the surface location. These plots (Figures \ref{fig:water-orientation}, \ref{fig:so2-orientation}, and \ref{fig:so2-transit-angles}) allow for a visual interpretation of how the net orientations are affected when moving from the gas phase through the interfacial region and to the surface location, and then further into the aqueous interfacial region and bulk. Both the neat-water, with only a single \suldiox~introduced, and the high-concentration saturated system were analyzed. In the case of the neat-water system, the introduction of a single \suldiox~does not greatly affect molecular orientation of water molecules in the interfacial region. These results of the water orientation are very similar to a neat-water system without any adsorbed solutes (not shown).

@@

Although the plots show overall similarities for both the neat-water and saturated systems, the presence of a layer of adsorbed \suldiox~molecules alters the orientation of those waters furthest into the gas phase. For the saturated solution, the resulting orientation of waters above 0\angs, shown in the bottom set of plots of Figure \ref{fig:water-orientation}, is nearly isotropic in $\phi$, and with $\theta \leq$ 90\textdegree. This results from waters with bisectors pointing further into the adsorbed \suldiox~gas layer, and both hydrogens pointing outward from the aqueous bulk. The effect is more pronounced as the waters move further from the water surface, and above 4\angs~the $\theta$ distribution is mostly concentrated around $\theta=$ 0\textdegree (see Figure \ref{fig:water-angles}).

@@

The distribution of $\phi$ is more sharply defined (i.e. less isotropic) for the neat-water system than for the saturated one. Waters on the neat surface lie flat or perpendicular to the surface if they are below or above the water surface location, respectively. The presence of the \suldiox~allows a greater range of ``twist'' for those waters in the plane of the interface. The $\phi$ distributions quickly become isotropic above the saturated water surface location, shown as a uniform coloration across $\phi$ for most values of $\theta$.

@@

Orientation distributions of the adsorbed \suldiox~molecules were created during the equilibrium simulations for both the neat-water and saturated systems. Figure \ref{fig:so2-orientation} shows the 2D distributions of $\theta$ and $\phi$ (arranged similarly to the water orientation distributions plots in Figure \ref{fig:water-orientation}). The \suldiox~orientation data set for the neat-water system is much smaller as only a single \suldiox~molecule was simulated in the bulk. The resulting distribution plots are thus representative of the single surface active \suldiox~molecule. The neat-water \suldiox~molecule remains within a narrower region of the interface than the saturated system \suldiox, but effective comparisons can still be drawn. Note that the depth range of the plots in Figure \ref{fig:so2-orientation} is different for the neat-water and saturated systems reflecting the surface mobility of the \suldiox~molecules in the two systems.

In the interfacial region the angular distribution of the single \suldiox~(in the neat-water system) is concentrated primarily in $\theta <$ 90 \textdegree. The peak of the distribution occurs at $\theta=$ 0\textdegree. This indicates that the \suldiox~bisector points out of the water surface, with the sulfur atom pointing towards the aqueous bulk, and the two oxygens pointing into the gas phase. This same distribution occurs in the saturated system for depths below the surface location, $< 0$\angs. Beyond 4\angs~above the surface, both distributions become mostly isotropic, either because the \suldiox~does not venture into the gas in the neat-water system, or because of the nature of the adsorbed \suldiox~layer in the saturated system. Promixity to the water surface highly orients the \suldiox~bisector. % with the sulfur atom pointing in towards the water bulk. %Moving further away from the water surface, and interacting less with \wat~molecules allows for greater orientational freedom as exhibited in the isotropy of the distributions above 4\angs. %Because both the neat-water and saturated systems show rather similar distributions, it is possible that the concentration of \suldiox~does not strongly affect the orientation of the \suldiox, unlike the orientations of waters.

The distributions of $\phi$ are isotropic in both systems at all depths. Because the \suldiox~bisector near the surface is oriented perpendicularly to the interface, the isotropy in $\phi$ is expected. Further from the water surface where the bisector orientation becomes isotropic, the $\phi$ distribution remains isotropic. For the surface \suldiox~orientation, the $\phi$ angle does not provide further information regarding the surface behavior or orientational preference.

@@

\begin{figure}[h!]
	\begin{center}
		\includegraphics[scale=1.0]{images/so2-angles/theta-phi.png}
		\caption{Molecular orientation distributions for \suldiox~molecules adsorbed to the water slab surface. The angular distributions are shown for the neat-water (top) and saturated (bottom) systems, and are arranged similarly to the water orientation plots of Figure \ref{fig:water-orientation}. For both systems the distributions show \suldiox~bound to the water surface with the sulfur primarily pointing towards the water slab, and the oxygens pointing to the gas phase. In this configuration the $\phi$ distribution is isotropic because of the water slab's in-plane symmetry. Note that the range of surface depths in the low concentration plot is half that of the saturated system. This is due to the narrower position range of the single \suldiox~molecule.}
		\label{fig:so2-orientation}
	\end{center}
\end{figure}

### 2011-10-25 Methods.tex
Polishing the review responses
Final resubmission to the journal

Molecular dynamics simulations were performed using the Amber 11 software suite.\cite{Case2010} Polarizable models for the \wat~and \suldiox~molecules were used in the simulations, and have been used previously in studies on interfacial systems because they are known to more accurately reproduce interfacial structure and free energy profiles.\cite{Wick2007,Rivera2006,Dang1998} The \wat~model used is the POL3 water model,\cite{Caldwell1995} and for \suldiox~we used the model of Baer et al. that places a single polarizable center on the sulfur atom.\cite{Baer2010} An intermolecular cutoff of 12\angs~was used for long-range electrostatic forces. The simulations were performed in the NVT ensemble using Langevin dynamics for temperature control. Induced dipoles were treated by the polarizable potential functions of the Amber molecular dynamics software.

@@

The low and high concentration simulated systems, ``neat-water'' and ``saturated water'', respectively, were created as follows: the neat-water simulation involved the addition of a single \suldiox~molecule either within the bulk of the water slab (for equilibrium MD), or above the slab surface (in the SMD simulations). The single-\suldiox~neat-water system was then evolved for 2 ns to produce an equilibrated starting configuration. Because of the extremely low concentration of \suldiox~in the neat-water system the surface waters behave similarly to a true neat-water air-liquid interface. The neat-water system orientational results shown later in this work reproduce well the results of our previous orientational studies of surface water behavior.\cite{Walker2006b,Hore2008} The saturated system had 22 \suldiox~molecules introduced to the water slab bulk in order to saturate it to a level coinciding with the Henry's law constant for \suldiox~in water (k\textdegree$_H = 1.4$ mol/kg*bar).\cite{Lide2000} Additionally, 50 \suldiox~molecules were introduced into the gas phase outside of the saturated water slab to simulate an added \suldiox~gas pressure. The saturated system with both bulk and gaseous \suldiox~was then evolved for 2 ns to produce a starting configuration for further saturated simulations.

@@

Figure \ref{fig:density-flaw} demonstrates the problem of surface location drift during a simulation. Figure \ref{fig:density-flaw}A shows the density profile of water and \suldiox~over the course of one of the 10 ns trajectories used in this work using the original uncorrected coordinates of the system taken from the raw atomic positions of the molecular dynamics data output. The water density profile and location (thin gray line in Figure \ref{fig:density-flaw}A) is produced by averaging the instantaneous density profile at each time step in the simulation over all the time steps. The water profile was then fit to a $\tanh$ function (black line) to extract the position and width parameters of the water surface. The fitted surface water density profile has a width of 3.77\angs, which is comparable to values reported for similar neat-water systems.\cite{Dang1997,Hore2008} A bulge in the gas-phase ($\leq 52$\angs) side of the water density profile is indicative of the drift of the water slab over the course of the trajectory. Thus the calculated location and width from the $\tanh$ line fit are not accurate over long trajectories for defining a stationary reference point.

@@

\begin{figure}[h!]
	\begin{center}
		\includegraphics[scale=1.0]{images/angle-cartoons/molecular-angles-small.png}
		\caption{(A) The definition of the angles $\theta$ and $\phi$ used to define molecular orientation of \suldiox~and \wat~relative to a reference axis. (B) $\theta$ is the value of the ``tilt'' of the molecular bisector from the reference axis, with $\theta$ values ranging from  0\textdegree~(aligned) to 180\textdegree~(anti-aligned). $\theta = $ 90\textdegree~places the bisector in the plane of the water surface. Several bisector orientations are shown depicting different $\theta$ values. (C) $\phi$ is the ``twist'' angle defined as a rotation around the molecular bisector axis. A value of 0\textdegree~aligns the plane of the molecule perpendicular to the plane of the water surface. A value of 90\textdegree~rotates the molecule to lie flat in the plane of the water surface. Because of the symmetries of \suldiox, \wat, and the plane of the water surface, the range of $\phi$ values is limited to 0\textdegree~$\leq \phi \leq$~90\textdegree.}
		\label{fig:water-angles}
	\end{center}
\end{figure}

>>>>

Molecular dynamics simulations were performed using the Amber 11 software suite.\cite{Case2010} Polarizable models for the \wat~and \suldiox~molecules were used in the simulations, and have been used previously in studies on interfacial systems because they are known to more accurately reproduce interfacial structure and free energy profiles.\cite{Wick2007,Rivera2006,Dang1998} The \wat~model used is the POL3 water model (also discussed in the supplemental information),\cite{Caldwell1995} and for \suldiox~we used the model of Baer et al. that places a single polarizable center on the sulfur atom.\cite{Baer2010} An intermolecular cutoff of 12\angs~was used for long-range electrostatic forces. The simulations were performed in the NVT ensemble using Langevin dynamics for temperature control. Induced dipoles were treated by the polarizable potential functions of the Amber molecular dynamics software.

@@

The low and high concentration simulated systems, ``neat-water'' and ``saturated'', respectively, were created as follows: the neat-water simulation involved the addition of a single \suldiox~molecule either within the bulk of the water slab (for equilibrium MD), or above the slab surface (in the SMD simulations). The single-\suldiox~neat-water system was then evolved for 2 ns to produce an equilibrated starting configuration. Because of the extremely low concentration of \suldiox~in the neat-water system the surface waters behave similarly to a true neat-water air-liquid interface. The neat-water system orientational results shown later in this work reproduce well the results of our previous orientational studies of surface water behavior.\cite{Walker2006b,Hore2008} The saturated system had 22 \suldiox~molecules introduced to the water slab bulk in order to saturate it to a level coinciding with the Henry's law constant for \suldiox~in water (k\textdegree$_H = 1.4$ mol/kg*bar).\cite{Lide2000} Additionally, 50 \suldiox~molecules were introduced into the gas phase outside of the saturated water slab to simulate an added \suldiox~gas pressure. The additional gas in the vapor phase was added over the course of several ns to keep a constant 1 atm of \suldiox~pressure above the water surface as \suldiox~gas molecules adsorbed to the surface. The saturated system with both bulk and gaseous \suldiox~was then evolved for 2 ns to produce a starting configuration for further saturated simulations.

@@

Figure \ref{fig:density-flaw} demonstrates the problem of surface location drift during a simulation (even with utilizing the Amber NSCM parameter). Figure \ref{fig:density-flaw}A shows the density profile of water and \suldiox~over the course of one of the 10 ns trajectories used in this work using the original uncorrected coordinates of the system taken from the raw atomic positions of the molecular dynamics data output. The water density profile and location (thin gray line in Figure \ref{fig:density-flaw}A) is produced by averaging the instantaneous density profile at each time step in the simulation over all the time steps. The water profile was then fit to a $\tanh$ function (black line) to extract the position and width parameters of the water surface. The fitted surface water density profile has a width of 3.77\angs, which is comparable to values reported for similar neat-water systems.\cite{Dang1997,Hore2008} A bulge in the gas-phase ($\leq 52$\angs) side of the water density profile is indicative of the drift of the water slab over the course of the trajectory. Thus the calculated location and width from the $\tanh$ line fit are not accurate over long trajectories for defining a stationary reference point.

@@

\begin{figure}[h!]
	\begin{center}
		\includegraphics[scale=1.0]{images/angle-cartoons/molecular-angles-small.png}
		\caption{(A) The definition of the angles $\theta$ and $\phi$ used to define molecular orientation of \suldiox~and \wat~relative to a reference axis. (B) $\theta$ is the value of the ``tilt'' of the molecular bisector from the reference axis, with $\theta$ values ranging from  0\textdegree~(aligned) to 180\textdegree~(anti-aligned). $\theta = $ 90\textdegree~aligns the bisector parallel to the plane of the water surface. Several bisector orientations are shown depicting different $\theta$ values. (C) $\phi$ is the ``twist'' angle defined as a rotation around the molecular bisector axis. A value of 0\textdegree~aligns the plane of the molecule perpendicular to the plane of the water surface. A value of 90\textdegree~rotates the molecule to lie flat in the plane of the water surface. Because of the symmetries of \suldiox, \wat, and the plane of the water surface, the range of $\phi$ values is limited to 0\textdegree~$\leq \phi \leq$~90\textdegree.}
		\label{fig:water-angles}
	\end{center}
\end{figure}
